---
title: "Integration of HB10hpf, HB13hpf, and HB16hpf neural subsets RNAseq_ATACseq samples R Notebook"
output: 
  html_notebook:
    toc: TRUE
    toc_depth: 2
    toc_float: TRUE
---

```{css, echo = F}
caption {
      color: blue;
      font-weight: bold;
      font-size: 2.0em;
    }
```

This integration strategy will be to use the ATAC-seq peaks from all 3 multiomic subset samples to create a peak set that will be used to add a Chromatin assay called "int_peaks" to each sample.  The multiomic sample can then be RNA-seq integrated which will merge the ATAC-seq assays. Will then use harmony to integrate the ATAC-seq based on the "int_peaks" assays and finally will integrate the Seurat integrated RNA-seq and harmony integrated ATAC-seq with weighted nearest neighbor for final cluster determination.

```{r, results=F}
library(Seurat)
library(org.Dr.eg.db)
library(BSgenome.Drerio.UCSC.danRer11)
library(Signac)
library(knitr)
library(kableExtra)
library(dplyr)
library(ggplot2)
library(ggsci)
library(limma)
library(JASPAR2020)
library(patchwork)
library(TFBSTools)
library(motifmatchr)
library(AnnotationHub)
library(harmony)
options(future.globals.maxSize = 4000 * 1024^2)
```

# Get annotation
```{r}
ah = AnnotationHub()
GRCz11.v99.EnsDb = ah[["AH78759"]]  ## /beevol/home/orourke/genomes/cellranger_arc_genomes/zebrafishPlusGFP built on GRCz11.v99.gtf
#GRCz11.v103.EnsDb <- ah[["AH89399"]]
```
```{r, results=FALSE, warning=FALSE}
# get gene annotations for GRCz11
annotation <- GetGRangesFromEnsDb(ensdb = GRCz11.v99.EnsDb, verbose = FALSE)
seqlevelsStyle(annotation) <- "UCSC"
genome(annotation) <- "GRCz11"
```

# Read data
```{r}
HB10hpf <- readRDS("~/Documents/Projects/Sagerstrom/scRNA-seq_ATAC-seq_reanalysis_2021_07_12/InitialExploration/Neural_subsets/RDSfiles/HB10hpf_neural.RDS")
HB13hpf <- readRDS("~/Documents/Projects/Sagerstrom/scRNA-seq_ATAC-seq_reanalysis_2021_07_12/InitialExploration/Neural_subsets/RDSfiles/HB13hpf_neural.RDS")
HB16hpf <- readRDS("~/Documents/Projects/Sagerstrom/scRNA-seq_ATAC-seq_reanalysis_2021_07_12/InitialExploration/Neural_subsets/RDSfiles/HB16hpf_neural.RDS")
```

# Common peak set
The first step in integration is to create a common ATAC peak set

```{r}
## note the peaks assay is the one created with the mac2 called peaks
HB10hpf@assays$peaks@ranges
HB13hpf@assays$peaks@ranges
HB16hpf@assays$peaks@ranges
```

```{r}
# Create a unified set of peaks to quantify in each dataset
combined.peaks <- reduce(x = c(HB10hpf@assays$peaks@ranges,HB13hpf@assays$peaks@ranges,HB16hpf@assays$peaks@ranges))
# Filter out bad peaks based on length
combined.peaks
peakwidths <- width(combined.peaks)
combined.peaks <- combined.peaks[peakwidths  < 10000 & peakwidths > 20]
combined.peaks
```
# Quantify peaks in each sample

The processed RNAseq/ATACseq samples already have a Fragment class object so can use that to create a new assay which will call int_peaks, to distinguish it from the original ATAC peaks assay from the 10X genomics peak calls and the peaks assay from the macs2 peak calls for the individual sample
```{r}
## first quantify combined peaks in each sample
DefaultAssay(HB10hpf) <- "peaks"
HB10hpf.counts <- FeatureMatrix(
  fragments = Fragments(HB10hpf),
  features = combined.peaks,
  cells = colnames(HB10hpf)
)
DefaultAssay(HB13hpf) <- "peaks"
HB13hpf.counts <- FeatureMatrix(
  fragments = Fragments(HB13hpf),
  features = combined.peaks,
  cells = colnames(HB13hpf)
)
DefaultAssay(HB16hpf) <- "peaks"
HB16hpf.counts <- FeatureMatrix(
  fragments = Fragments(HB16hpf),
  features = combined.peaks,
  cells = colnames(HB16hpf)
)
```

```{r}
## next create chromatin assay for each sample
HB10hpf[["int_peaks"]] <- CreateChromatinAssay(
    counts = HB10hpf.counts,
    fragments = Fragments(HB10hpf),
    annotation = annotation,
    genome = 'GRCz11'
  )
HB13hpf[["int_peaks"]] <- CreateChromatinAssay(
    counts = HB13hpf.counts,
    fragments = Fragments(HB13hpf),
    annotation = annotation,
    genome = 'GRCz11'
  )
HB16hpf[["int_peaks"]] <- CreateChromatinAssay(
    counts = HB16hpf.counts,
    fragments = Fragments(HB10hpf),
    annotation = annotation,
    genome = 'GRCz11'
  )
```

```{r}
HB10hpf <- FindTopFeatures(HB10hpf, min.cutoff = 10)
HB10hpf <- RunTFIDF(HB10hpf)
HB10hpf <- RunSVD(HB10hpf)

HB13hpf <- FindTopFeatures(HB13hpf, min.cutoff = 10)
HB13hpf <- RunTFIDF(HB13hpf)
HB13hpf <- RunSVD(HB13hpf)

HB16hpf <- FindTopFeatures(HB16hpf, min.cutoff = 10)
HB16hpf <- RunTFIDF(HB16hpf)
HB16hpf <- RunSVD(HB16hpf)
```

Add orig.ident info to each
```{r}
HB10hpf$orig.ident <- "HB10hpf"
HB13hpf$orig.ident <- "HB13hpf"
HB16hpf$orig.ident <- "HB16hpf"
```

```{r}
saveRDS(HB10hpf, file = "RDSfiles/HB10hpf.neural.int_peaks.assay.RDS")
saveRDS(HB13hpf, file = "RDSfiles/HB13hpf.neural.int_peaks.assay.RDS")
saveRDS(HB16hpf, file = "RDSfiles/HB16hpf.neural.int_peaks.assay.RDS")
```

Integrate samples by RNA-seq (will also merge ATAC-seq)

```{r}
RNA_integration <- function(seurat.list){
  features <- SelectIntegrationFeatures(object.list = seurat.list, nfeatures = 3000)
  seurat.list <- PrepSCTIntegration(object.list = seurat.list, anchor.features = features)
  seurat.anchors <- FindIntegrationAnchors(object.list = seurat.list, normalization.method = "SCT", 
                                           anchor.features = features)
  seurat.int <- IntegrateData(anchorset = seurat.anchors, normalization.method = "SCT")
  seurat.int <- RunPCA(seurat.int, verbose = FALSE)
  seurat.int <- RunUMAP(seurat.int, dims = 1:50, reduction.name = 'umap.rna', reduction.key = 'rnaUMAP_')
  
  return(seurat.int)
}
```

```{r}
DefaultAssay(HB10hpf) <- "SCT"
DefaultAssay(HB13hpf) <- "SCT"
DefaultAssay(HB16hpf) <- "SCT"
HB10hpf$orig.ident <- "HB10hpf"
HB13hpf$orig.ident <- "HB13hpf"
HB16hpf$orig.ident <- "HB16hpf"
```

# integration 3WT
```{r}
list.3WT <- list(HB10hpf,HB13hpf,HB16hpf)
int.3WT <- RNA_integration(list.3WT)
```

```{r}
int.3WT[["int_peaks"]]
```

```{r}
saveRDS(int.3WT, file = "RDSfiles/int.neural.3WT.RDS")
```

# test of simple merge of ATAC seq

```{r}
GetLSI <- function(seurat){
  seurat <- FindTopFeatures(seurat, min.cutoff = 10)
  seurat <- RunTFIDF(seurat)
  seurat <- RunSVD(seurat)
  seurat <- RunUMAP(seurat, reduction = 'lsi', dims = 2:50)
  
  return(seurat)
}
```

## int.3WT
```{r}
DefaultAssay(int.3WT) <- "int_peaks"
int.3WT <- GetLSI(int.3WT)
p1.int.3WT <- DimPlot(int.3WT, group.by = "orig.ident")
p1.int.3WT
```

# Harmony integration of ATAC-seq
```{r}
RunATACharmony <- function(seurat){
  seurat <- RunHarmony(object=seurat,group.by.vars='orig.ident',
                                  reduction='lsi',assay.use='int_peaks',project.dim=F)
  seurat <- RunUMAP(seurat,dims=2:50,reduction='harmony',reduction.name = "umap.atac",reduction.key = "atacUMAP_")
  seurat <- FindNeighbors(seurat, dims = 1:50,reduction='harmony')
  seurat <- FindClusters(seurat, resolution = 2,reduction='harmony')
  seurat <- BuildClusterTree(object = seurat,reorder = T,reorder.numeric = T,verbose = T)
  
  return(seurat)
}
```

## int.3WT
```{r, fig.width=10}
DefaultAssay(int.3WT) <- "int_peaks"
int.3WT <- RunATACharmony(int.3WT)
p2.int.3WT <- DimPlot(int.3WT,group.by='orig.ident',pt.size=0.1, reduction = "umap.atac") + ggplot2::ggtitle('int.3WT')
p3.int.3WT <- DimPlot(int.3WT, group.by = "int_peaks_snn_res.2",pt.size=0.1, reduction = "umap.atac") + scale_color_igv()
p2.int.3WT + p3.int.3WT
```

```{r, fig.width=10, fig.height=3}
(p1.int.3WT + ggtitle("int.3WT Merged")) | (p2.int.3WT + ggtitle("int.3WT Integrated"))
```

```{r}
saveRDS(int.3WT, file = "RDSfiles/int.neural.3WT.RDS")
```

# Weighted Nearest Neighbor integratio of RNAseq and ATACseq

```{r, results=F}
RunWnnRnaAtac <- function(seurat){
  seurat <- FindMultiModalNeighbors(seurat, reduction.list = list("pca", "harmony"), dims.list = list(1:50, 2:50))
  seurat <- RunUMAP(seurat, nn.name = "weighted.nn", reduction.name = "wnn.umap", reduction.key = "wnnUMAP_", assay = "RNA")
  seurat <- FindClusters(seurat, graph.name = "wsnn", algorithm = 3, verbose = FALSE)
}

```

## int.3WT
```{r}
DefaultAssay(int.3WT) <- "integrated"
int.3WT <- RunWnnRnaAtac(int.3WT)
```

```{r, fig.width=10, fig.height=5, results=F}
p.3WT.rna <- DimPlot(int.3WT, reduction = "umap.rna", label = TRUE, label.size = 5, repel = TRUE) + ggtitle("integrated RNA")
p.3WT.atac <- DimPlot(int.3WT, reduction = "umap.atac", label = TRUE, label.size = 5, repel = TRUE) + ggtitle("integrated ATAC")
p.3WT.wnn <- DimPlot(int.3WT, reduction = "wnn.umap", label = TRUE, label.size = 5, repel = TRUE) + ggtitle("integrated WNN")
p.3WT.rna + p.3WT.atac + p.3WT.wnn & NoLegend() & theme(plot.title = element_text(hjust = 0.5))
p.3WT.rna
p.3WT.atac
p.3WT.wnn
```

```{r, fig.width=10, fig.height=5, results=F}
p.3WT.rna <- DimPlot(int.3WT, reduction = "umap.rna",group.by = "orig.ident") + ggtitle("integrated RNA") + NoLegend()
p.3WT.atac <- DimPlot(int.3WT, reduction = "umap.atac",group.by = "orig.ident") + ggtitle("integrated ATAC") + NoLegend()
p.3WT.wnn <- DimPlot(int.3WT, reduction = "wnn.umap",group.by = "orig.ident") + ggtitle("integrated WNN")
p.3WT.rna + p.3WT.atac + p.3WT.wnn & theme(plot.title = element_text(hjust = 0.5))
```



```{r}
saveRDS(int.3WT, file = "RDSfiles/int.neural.3WT.RDS")
```


```{r}
sessionInfo()
```


