---
title: "Preprocessing and QC of scRNAseq/ATACseq samples R Notebook"
output:   
  html_notebook:
    toc: TRUE
    toc_depth: 3
    toc_float: TRUE
---
```{css, echo = F}
caption {
      color: blue;
      font-weight: bold;
      font-size: 2.0em;
    }
```



# 1. Libraries, annotation and palette
```{r, results=F}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr)
  library(org.Dr.eg.db)
  library(BSgenome.Drerio.UCSC.danRer11)
  library(Signac)
  library(ggplot2)
  library(ggsci)
  library(limma)
  library(AnnotationHub)
  library(reticulate)
  library(patchwork)
  use_python("/Users/becky/opt/miniconda3/bin/python")
  library(JASPAR2020)
  library(TFBSTools)
  library(motifmatchr)
}) 
options(future.globals.maxSize = 4000 * 1024^2)
```

```{r}
ah = AnnotationHub()
GRCz11.v99.EnsDb = ah[["AH78759"]]  ## /beevol/home/orourke/genomes/cellranger_arc_genomes/zebrafishPlusGFP built on GRCz11.v99.gtf
#GRCz11.v103.EnsDb <- ah[["AH89399"]]
```

```{r}
mypal <- pal_igv(palette = "default",alpha = 1)(30)
```

# 2. Functions

## 2.1 Create Seurat object
```{r, results=F}
SeuratFromMatrixh5 <- function(inputdata, annotation, frag){
  # the 10x hdf5 file contains both data types
  rna_counts <- inputdata$`Gene Expression`
  atac_counts <- inputdata$Peaks
  
  seurat <- CreateSeuratObject(counts = rna_counts, assay = "RNA")
  seurat[["percent.mt"]] <- PercentageFeatureSet(seurat, pattern = "^mt-")
  
  seurat[["ATAC"]] <- CreateChromatinAssay(
    counts = atac_counts,
    sep = c(":", "-"),
    fragments = frag,
    annotation = annotation)
  
  return(seurat)
}
```

## 2.2 QC plots
```{r, fig.height=5, fig.width=7  }
GetQCplots <- function(seurat, mitoPattern){
  seurat[["percent.mt"]] <- PercentageFeatureSet(seurat, pattern = mitoPattern)
  v <- VlnPlot(seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
  f1 <- FeatureScatter(seurat, feature1 = "nCount_RNA", feature2 = "percent.mt")
  f2 <- FeatureScatter(seurat, feature1 = "nFeature_RNA", feature2 = "percent.mt")
  f3 <- FeatureScatter(seurat, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
  
  qc1 <- ggplot(seurat@meta.data, aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mt)) + geom_point(size=0.1) +
    scale_color_gradient(low="blue",high="red") + theme_classic()

  qc2 <- ggplot(seurat@meta.data, aes(x=nCount_RNA, y=percent.mt)) + geom_point(size=0.1) + scale_x_log10() +
    geom_density2d()

  qc3 <- ggplot(seurat@meta.data, aes(x=nCount_RNA, y=nFeature_RNA ,color=percent.mt)) + geom_point(size=0.1) +
    scale_x_log10() + scale_y_log10() + geom_density2d() +
    scale_color_gradient(low="gray",high="darkblue") + theme_classic()

  print(v)
  print(f1)
  print(f2)
  print(f3)
  print(qc1)
  print(qc2)
  print(qc3)

  return(seurat)
}
```

```{r}
GetATACplots <- function(seurat){
  DefaultAssay(seurat) <- "ATAC"
  
  seurat <- NucleosomeSignal(seurat)
  seurat <- TSSEnrichment(seurat)
  
  v1 <- VlnPlot(
    object = seurat,
    features = "nCount_ATAC",
    pt.size = 0
  ) + NoLegend()
  v2 <- VlnPlot(
    object = seurat,
    features = "nCount_ATAC",
    y.max = 100000,
    pt.size = 0
  ) + NoLegend()
  v3 <- VlnPlot(
    object = seurat,
    features = "nCount_ATAC",
    y.max = 5000,
    pt.size = 0
  ) + NoLegend()
  
  v4 <- VlnPlot(
    object = seurat,
    features = "TSS.enrichment",
    pt.size = 0
  ) + NoLegend()
  v5 <- VlnPlot(
    object = seurat,
    features = "TSS.enrichment",
    y.max = 10,
    pt.size = 0
  ) + NoLegend()
  
  v6 <- VlnPlot(
    object = seurat,
    features = "nucleosome_signal",
    pt.size = 0
  ) + NoLegend()
  v7 <- VlnPlot(
    object = seurat,
    features = "nucleosome_signal",
    y.max = 2,
    pt.size = 0
  ) + NoLegend()

  print(v1 + v2 + v3 + plot_layout(ncol = 3))
  print(v4 + v5 + plot_layout(ncol = 2))
  print(v6 + v7 + plot_layout(ncol = 2))
  
  return(seurat)
}
```

## 2.3 Call Peaks with macs2 and filter
```{r}
CreatePeaksWithFilter <- function(seurat, frag, blacklist, annotation){
  peaks <- CallPeaks(frag, macs2.path = "/Users/becky/opt/miniconda3/bin/macs2")
  len.ATAC <- length(seurat@assays$ATAC@ranges)
  len.peaks <- length(peaks)
  
  peaks <- keepStandardChromosomes(peaks, pruning.mode = "coarse")
  len.std <- length(peaks)
  peaks <- subsetByOverlaps(x = peaks, ranges = blacklist, invert = TRUE)
  len.bl <- length(peaks)
  ChrEnds <- GRanges(seqnames = annotation@seqinfo@seqnames,
                     ranges = IRanges(start = annotation@seqinfo@seqlengths - 300,
                                      end = annotation@seqinfo@seqlengths))
  ChrStarts <- GRanges(seqnames = annotation@seqinfo@seqnames,
                       ranges = IRanges(start = rep(1,26), end = rep(300,26)))
  RemoveChrEndsChrStarts <- c(ChrEnds,ChrStarts)
  peaks <- subsetByOverlaps(x = peaks, ranges = RemoveChrEndsChrStarts, invert = TRUE)
  len.rends <- length(peaks)
  
  cat("Number original ATAC-seq peaks: ",len.ATAC,"\n")
  cat("Number macs2 called ATAC-seq peaks: ",len.peaks,"\n")
  cat("Number macs2 ATAC-seq peaks with Standard Chromosomes: ",len.std,"\n")
  cat("Number macs2 ATAC-seq peaks after blacklist filter: ",len.bl,"\n")
  cat("Number macs2 ATAC-seq peaks after 300bp end filter: ",len.rends,"\n")
  
  macs2_counts <- FeatureMatrix(
    fragments = Fragments(seurat),
    features = peaks,
    cells = colnames(seurat)
  )
  
  seurat[["peaks"]] <- CreateChromatinAssay(
    counts = macs2_counts,
    fragments = frag,
    annotation = annotation,
    genome = 'GRCz11'
  )
  
  DefaultAssay(seurat) <- "peaks"
  
  seurat <- NucleosomeSignal(seurat)
  seurat <- TSSEnrichment(seurat)
  
  return(seurat)
  
}

```

## 2.4 UMAPS and clusters
```{r, results=F, warnings=F,fig.width=15, fig.height=5}
GetUMAPandClusters <- function(seurat){
  # RNA analysis
  DefaultAssay(seurat) <- "RNA"
  seurat <- SCTransform(seurat, verbose = FALSE, return.only.var.genes = FALSE) 
  seurat <- RunPCA(seurat) 
  ElbowPlot(seurat,ndims = 50)
  seurat <- RunUMAP(seurat, dims = 1:50, reduction.name = 'umap.rna', reduction.key = 'rnaUMAP_')
  
  # ATAC analysis
  # We exclude the first dimension as this is typically correlated with sequencing depth
  DefaultAssay(seurat) <- "peaks"
  
  seurat <- FindTopFeatures(seurat, min.cutoff = 5)
  seurat <- RunTFIDF(seurat)
  seurat <- RunSVD(seurat)
  seurat <- RunUMAP(seurat, reduction = 'lsi', dims = 2:50, reduction.name = "umap.atac", reduction.key = "atacUMAP_")
  
  DefaultAssay(seurat) <- "SCT"
  seurat <- FindMultiModalNeighbors(seurat, reduction.list = list("pca", "lsi"), dims.list = list(1:50, 2:50))
  seurat <- RunUMAP(seurat, nn.name = "weighted.nn", reduction.name = "wnn.umap", reduction.key = "wnnUMAP_", assay = "RNA")
  seurat <- FindClusters(seurat, graph.name = "wsnn", algorithm = 3, verbose = FALSE)
  seurat <- FindClusters(seurat, graph.name = "wsnn", algorithm = 3, verbose = FALSE, resolution = 1.0)
seurat <- FindClusters(seurat, graph.name = "wsnn", algorithm = 3, verbose = FALSE, resolution = 1.5)
seurat <- FindClusters(seurat, graph.name = "wsnn", algorithm = 3, verbose = FALSE, resolution = 2.0)
seurat <- FindClusters(seurat, graph.name = "wsnn", algorithm = 3, verbose = FALSE, resolution = 2.5)
seurat <- FindClusters(seurat, graph.name = "wsnn", algorithm = 3, verbose = FALSE, resolution = 3.0)
seurat <- FindClusters(seurat, graph.name = "wsnn", algorithm = 3, verbose = FALSE, resolution = 3.5)
  
  p1 <- DimPlot(seurat, reduction = "umap.rna", label = TRUE, label.size = 5, repel = TRUE, group.by = "wsnn_res.0.8") + ggtitle("RNA") + scale_color_igv()
  p2 <- DimPlot(seurat, reduction = "umap.atac", label = TRUE, label.size = 5, repel = TRUE, group.by = "wsnn_res.0.8") + ggtitle("ATAC") + scale_color_igv()
  p3 <- DimPlot(seurat, reduction = "wnn.umap", label = TRUE, label.size = 5, repel = TRUE, group.by = "wsnn_res.0.8") + ggtitle("WNN") + scale_color_igv()
  print(p1 + p2 + p3 & NoLegend() & theme(plot.title = element_text(hjust = 0.5)))
  
  return(seurat)
}
```

# 3 Read in data
Use same annotation for each sample
```{r, results=FALSE, warning=FALSE}
# get gene annotations for GRCz11
annotation <- GetGRangesFromEnsDb(ensdb = GRCz11.v99.EnsDb, verbose = FALSE)
seqlevelsStyle(annotation) <- "UCSC"
genome(annotation) <- "GRCz11"
```

## 3.1 HB13hpf.rep2
```{r, results=F}
inputdata <- Read10X_h5("../data/HB13hpf_rep2/filtered_feature_bc_matrix.h5")
frag.file <- "../data/HB13hpf_rep2/atac_fragments.tsv.gz"
HB13hpf.rep2 <- SeuratFromMatrixh5(inputdata, annotation, frag.file)
```

```{r}
HB13hpf.rep2
```

# 4 QC Plots

## 4.1 RNA-seq QC plots

### 4.1.1 HB13hpf.rep2
```{r}
HB13hpf.rep2 <- GetQCplots(HB13hpf.rep2, "^mt-")
```


## 4.2 ATAC-seq QC plots

### 4.2.1 HB13hpf.rep2
```{r}
HB13hpf.rep2 <- GetATACplots(HB13hpf.rep2)
```

# 5 Filter samples

## 5.1 HB13hpf.rep2

It's odd that the higher percent.mt cells are also higher total RNA count cells.  Will filter at <20% percent.mt to keep these and see what they are.
```{r}
cells.orig <- table(HB13hpf.rep2@meta.data$orig.ident)[[1]]
HB13hpf.rep2 <- subset(x = HB13hpf.rep2,
  subset = nCount_ATAC < 50000 & nCount_ATAC > 700)
cells.atac <- table(HB13hpf.rep2@meta.data$orig.ident)[[1]]
HB13hpf.rep2 <- subset(x = HB13hpf.rep2,
  subset = nFeature_RNA < 4000 & nFeature_RNA > 500)
cells.rna.f <- table(HB13hpf.rep2@meta.data$orig.ident)[[1]]
HB13hpf.rep2 <- subset(x = HB13hpf.rep2,
  subset =percent.mt < 10)
cells.mt <- table(HB13hpf.rep2@meta.data$orig.ident)[[1]]
HB13hpf.rep2 <- subset(x = HB13hpf.rep2,
  subset = nucleosome_signal < 2 )
cells.nuc <- table(HB13hpf.rep2@meta.data$orig.ident)[[1]]
HB13hpf.rep2 <- subset(x = HB13hpf.rep2,
  subset = TSS.enrichment >1)
cells.tss <- table(HB13hpf.rep2@meta.data$orig.ident)[[1]]
cat("HB13hpf.rep2:","\n")
cat("cells in original sample: ",cells.orig,"\n")
cat("cells after ATAC < 50000 & ATAC > 700: ",cells.atac,"\n")
cat("cells after features RNA < 4000 and RNA > 500: ",cells.rna.f,"\n")
cat("cells after percent.mt < 10: ",cells.mt,"\n")
cat("cells after nucleosome < 2: ",cells.nuc,"\n")
cat("cells after TSS enrichment > 1",cells.tss,"\n")
```


# 6 Peak calling
Per recommendation from Signac tutorial call ATAC peaks with macs2 and put in assay "peaks"

#### GRCz11.blacklist
Found GRCz10 blacklist in Yang, H., Luan, Y., Liu, T. et al. A map of cis-regulatory elements and 3D genome structures in zebrafish. Nature 588, 337–343 (2020). https://doi.org/10.1038/s41586-020-2962-9
Supplementary table 19: Supplemental Table 19. ChIP-seq black list in the zebrafish genome.xlsx
```{r}
### code used to generate blacklist
# library(readxl)
# library(GenomicRanges)
# library(plyranges)
# GRCz10.blacklist <- read_excel("Supplemental_Table_19_ChIPseq_black_list_in_the_zebrafish_genome.xlsx", col_names = F)
# names(GRCz10.blacklist) <- c("seqnames","start","end")
# head(GRCz10.blacklist)
# GRCz10.blacklist <- as_granges(GRCz10.blacklist)
# head(GRCz10.blacklist)
# chain <- import.chain("danRer10ToDanRer11.over.chain")
# seqlevelsStyle(GRCz10.blacklist) = "UCSC"  # necessary
# GRCz11.blacklist = liftOver(GRCz10.blacklist, chain)
# class(GRCz11.blacklist)
# GRCz11.blacklist = unlist(GRCz11.blacklist)
# genome(GRCz11.blacklist) = "GRCz11"
# GRCz11.blacklist
# saveRDS(GRCz11.blacklist, file = "GRCz11.blacklist.granges.RDS")
```

```{r}
GRCz11.blacklist <- readRDS(file = "~/Documents/Projects/R_General/GRCz11.blacklist.granges.RDS")
```

Call peaks with macs2 and filter with blacklist and filter out peaks in 300bp start and end of each chromosome which causes errors in Signac in downstream analyses

## 6.1 HB13hpf.rep2
```{r}
HB13hpf.rep2 <- CreatePeaksWithFilter(HB13hpf.rep2, frag.file, GRCz11.blacklist, annotation)
```

# 7 UMAPs

## 7.1 HB13hpf.rep2
```{r, results = F, fig.height=5, fig.width=15}
HB13hpf.rep2 <- GetUMAPandClusters(HB13hpf.rep2)
```

```{r, fig.height=5, fig.width=15}
(FeaturePlot(HB13hpf.rep2, features = "nFeature_RNA", reduction = "wnn.umap", max.cutoff = 2000)) + (FeaturePlot(HB13hpf.rep2, features = "percent.mt", reduction = "wnn.umap")) + (FeaturePlot(HB13hpf.rep2, features = "nFeature_ATAC", reduction = "wnn.umap", max.cutoff = 10000)) + plot_layout(ncol = 3)
```

# 8 Add Motifs

#### JASPAR2020 motif matrix
```{r}
# Get a list of motif position frequency matrices from the JASPAR database
pfm <- getMatrixSet(
  x = JASPAR2020,
  opts = list(collection="CORE", all_versions = FALSE)
)
```

## 8.1 HB13hpf.rep2
```{r, results=F}
DefaultAssay(HB13hpf.rep2) <- "peaks"
HB13hpf.rep2 <- RegionStats(HB13hpf.rep2, genome = BSgenome.Drerio.UCSC.danRer11, assay = "peaks")

HB13hpf.rep2 <- AddMotifs(
  object = HB13hpf.rep2,
  genome = BSgenome.Drerio.UCSC.danRer11,
  pfm = pfm
)
```

# 9 Save processed files

```{r}
if (!dir.exists("RDSfiles")){
  dir.create("RDSfiles")
}else{
  print("dir exists")
}
```

```{r}
saveRDS(HB13hpf.rep2, file = "RDSfiles/HB13hpf_rep2.clustered.RDS")
```

```{r}
sessionInfo()
```

